Natural Language Processing\\ Fall 2015 Final Project:\\ Sense Disambiguation with the\\ Naive Bayes Classifier

### 2015-12-13 paper/paper.tex
Mostly finish algorithm explanation

\documentclass[11pt]{article}
\usepackage{acl2015, latexsym, url, times, amssymb, amsmath, amsfonts}

@@

There are many approaches to word sense disambiguation — in particular machine learning tasks can be broadly adapted to solve the problem. One of the simpler and more efficient of such methods is the Naive Bayes Classifier, which is a probabilistic method that manages to achieve good results while still remaining fairly simple in implementation.

\section{The Naive Bayes Classifier}
Explain in exact detail the Naïve Bayes Classifier algorithm. This should more or less follow the format of project proposal

>>>>

\documentclass[11pt]{article}
\usepackage{acl2015, latexsym, url, times, amssymb, amsmath, amsfonts, mathtools}

\DeclareMathOperator*{\argmax}{argmax}

@@

There are many approaches to word sense disambiguation---in particular, machine learning tasks can be broadly adapted to solve the problem. One of the simpler and more efficient of such methods is the Naive Bayes Classifier, which is a probabilistic method that manages to achieve good results while still remaining fairly simple in implementation.

\section{The Naive Bayes Classifier}
As mentioned previously, humans are highly proficient at identifying the sense of a word in use when provided the context of that word. We identify contextual cues that tell us what a specific word's intended meaning must be. We will call these cues \emph{features}, and they may include anything that allows a reader to narrow down the choices for possible meanings. Two of the most important kinds of features are those that provide local context and those that provide global context. Local context refers to looking at how the word sits in the sentence, namely what words it is neighboring, their parts of speech, etc. Global context is the larger semantic space that the word exists in, like the topic of the sentence, paragraph, or text as whole. Every word in a given corpus can be assigned a vector of features providing this context, which we will denote $F$.

Given a set $F$ for each word $w$, we can then view the problem of disambiguating word sense from a probabilistic perspective. For all $w$, we are then attempting to choose the sense that is maximally probable, $s^*$, given $w$ and the observed feature set $F(s)$ for all words with sense $s$. That is:

\begin{equation} \label{eq:fundamental}
\begin{split}
s^*(w) &= \argmax_{s \in S(w)} P(s|F(s), w) \\
       &= \argmax_{s \in S(w)} \frac{P(F(s)|s,w)P(s|w)}{P(F(s))}
\end{split}
\end{equation}

This equation is simple and conceptually effective. But in practice, even the second form is difficult to use---to compute the probability that a given feature vector will occur requires that that entire feature vector have been encountered for that sense before. For large feature vectors the likelihood of this is extremely low, and requires prohibitively large sets of training data. The Naive Bayes Classifier (NBC) is an attempt to minimize the combinatorics of the features by making the \emph{na\"ive} assumption that the occurrence of each feature is an independent event. With this assumption, we may say that ${P(F(s)|s) = \prod_{f \in F(s)} P(f|s)}$. This assumption is certainly not always true, and is what gives the NBC its na\"ive qualifier. Indeed this equation is often an approximation at best, but usually it is sufficiently accurate to be useable.

Substituting this value in and removing $P(F(s))$, as it is constant for all $s$ and will not affect which is chosen for any given word, we arrive at the Naive Bayes Classifier equation:

\begin{equation} \label{eq:nbc}
s^*(w) = \argmax_{s \in S(w)} P(s|w)\!\!\prod_{f \in F(s)}\!\!P(f|s)
\end{equation}

With this formulation, the primary problem in effectively disambiguating word senses is one of training. Identifying the features that accurately determine word sense and finding sense-tagged corpora that exhibit those features are the primary problems in applying the NBC.
